---
title: "Programming in Base R"
format: html
editor: visual
---

## Task 1: Basic Vector Practice

Create vectors.

```{r}
prebp <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)

postbp <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)
```

Assign names to the vectors.

```{r}
names(prebp) <- paste("Subject", 1:20, sep = "_")
names(postbp) <- paste("Subject", 1:20, sep = "_")
```

Calculating the change in blood pressure between measurements.

```{r}
diffbp <- prebp-postbp
```

Calculating average decrease in blood pressure across all patients.

```{r}
mean(diffbp)
```

Determine which patients experienced a decrease in blood pressure after treatment (a positive change).

```{r}
which(diffbp > 0)
```

Subset differences vector to only return positive change.

```{r}
positivechange <- diffbp > 0
positivechange2 <- diffbp[positivechange]
```

Calculating average decrease in blood pressure for those where the blood pressure decreased (positive change).

```{r}
mean(positivechange2)
```

## Task 2: Basic Data Frame Practice

Creating data frame.

```{r}
bp_df <- data.frame(patient = names(prebp), pre_bp = prebp, post_bp = postbp, diff_bp = diffbp)
```

Return rows where diffpb column is negative.

```{r}
final_bp_df <- subset(bp_df, diffbp < 0)
```

Adding column where post_bp is less than 120.

```{r}
final_bp_df$post_bp_less_than_120 <- final_bp_df$post_bp < 120
```

Printing the data frame.

```{r}
knitr::kable(final_bp_df)
```

## Task 3: List Practice

Creating vectors with the new placebo data set.

```{r}
prebp2 <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
postbp2 <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
```

Naming vectors

```{r}
names(prebp2) <- paste("Subject", 1:10, sep = "_")
names(postbp2) <- paste("Subject", 1:10, sep = "_")
```

Calculating change in blood pressure.

```{r}
diffbp2 <- prebp2-postbp2
```

Determine which patients experienced a decrease in blood pressure after treatment (a positive change).

```{r}
which(diffbp2 > 0)
```

Subset differences vector to only return positive change.

```{r}
positivechange01 <- diffbp > 0
positivechange02 <- diffbp[positivechange]
```

Creating a data frame from placebo data.

```{r}
bp_df2 <- data.frame(patient = names(prebp2), pre_bp = prebp2, post_bp = postbp2, diff_bp = diffbp2)
```

Returning postbp values of less than 120.

```{r}
bp_df2$post_bp_less_than_120 <- postbp2 < 120
```

Returning rows where diffbp is negative

```{r}
final_bp_df2 <- subset(bp_df2, diffbp2 < 0)
```

Creating a list from the two data frames.

```{r}
bplist <- list(treatment = final_bp_df, placebo = final_bp_df2)
```

Accessing the first list element using three types of syntax.

```{r}
bplist [1]
```

```{r}
bplist [[1]]
```

```{r}
bplist$treatment
```

Accessing prebp column of placebo data frame.

```{r}
head(bplist$placebo[2])
```
